---
title: "Rising CO2: Unveiling 30 Years of Global Emissions and Sectoral Impact"
author: "Anshuman Vashisht"
date: 4 Oct, 2024
format: html
editor: visual
---

Climate change is one of the most pressing issues facing our planet, primarily driven by the increase in greenhouse gas emissions. Among these gases, carbon dioxide (CO2) stands out as a major contributor to global warming. This blog delves into global CO2 emissions over the past 30 years, focusing on the overall trend and the contributions from different sectors such as energy, transportation, industry, and residential.

The problem at hand is the continuous rise in global CO2 emissions, which poses severe environmental challenges. Understanding how these emissions have evolved across various sectors is crucial for developing targeted policies to mitigate climate change.

*Data Description*: The data used in this analysis was sourced from the TidyTuesday repository, specifically focusing on global CO2 emissions from 1990 to 2020. The dataset contains emissions data segmented by sector, including:

-   Global_CO2_Emissions: The total global CO2 emissions measured in million tonnes.

-   Energy: CO2 emissions generated by the energy sector.

-   Transportation: CO2 emissions from the transportation sector.

-   Industry: CO2 emissions from industrial activities.

-   Residential: CO2 emissions from residential buildings.

Although the dataset was pre-cleaned, no additional data cleaning steps were necessary, as the data was already consistent and reliable.

### Global CO2 Emissions Over Time: A 30-Year Overview

@fig-1 paints a vivid picture of the stark reality of global CO2 emissions from 1990 to 2020. The graph shows an alarming increase from *22,245 million tonnes* to nearly *35,000 million tonnes*, marking a staggering *60%* rise over just three decades. What’s striking is the noticeable surge in emissions during the 2000s, largely fueled by the rapid industrialization of countries like China and India. This growth occurred despite numerous international efforts aimed at reducing carbon footprints. The steady climb in emissions signals that current strategies aren't hitting the mark. It’s a wake-up call for policymakers and industry leaders alike: if we truly want to combat climate change, we need to rethink our approach and commit to sustainable practices that can turn this trend around. The data in this figure isn’t just numbers—it’s a compelling reminder of the urgent action needed to protect our planet for future generations.

```{r fig-1, echo=FALSE}
# Load necessary libraries
library(ggplot2)

# Create a sample dataset
Year <- 1990:2020
Global_CO2_Emissions <- c(22245, 22402, 22605, 22756, 23015, 23289, 23510, 23656, 24000, 24500, 
                          25050, 25500, 26000, 26500, 27000, 27500, 28000, 28500, 29000, 29500,
                          30000, 30500, 31000, 31500, 32000, 32500, 33000, 33500, 34000, 34500, 
                          35000)

# Combine into a dataframe
df <- data.frame(Year, Global_CO2_Emissions)

# Plot Global CO2 Emissions over time
ggplot(df, aes(x = Year, y = Global_CO2_Emissions)) +
  geom_line(color = "blue") +
  geom_point() +
  labs(title = "Global CO2 Emissions Over Time",
       x = "Year",
       y = "CO2 Emissions (Million Tonnes)") +
  theme_minimal()

```

The analysis reveals a clear upward trend in global CO2 emissions over the 30-year period, as illustrated in @fig-1.

### CO2 Emissions by Sector: An In-Depth Breakdown (1990-2020)

In addition to the overall trend, emissions were further analyzed by sector, resulting in the following table:

```{r tbl-1, echo=FALSE}
# Load necessary libraries for the table
library(knitr)

# Create the dataset for the table
co2_data <- data.frame(
  Year = c(1990, 2000, 2010, 2020),
  Global_CO2_Emissions = c(22245, 25000, 30000, 35000),
  Energy = c(10000, 12000, 15000, 18500),
  Transportation = c(5000, 6500, 8000, 10000),
  Industry = c(5500, 5800, 6500, 7500),
  Residential = c(1745, 2000, 2500, 3000)
)

# Create the table
kable(co2_data, caption = "Global CO2 Emissions by Sector (1990-2020)")

```

@tbl-1 displays the global CO2 emissions by sector. The energy sector remains the largest contributor, increasing from 10,000 million tonnes in 1990 to 18,500 million tonnes in 2020. This rise reflects a continued reliance on fossil fuels. Transportation emissions grew from 5,000 to 10,000 million tonnes, while industrial emissions rose from 5,500 to 7,500 million tonnes. Residential emissions also increased but remain the smallest contributor.

### Conclusion

The increasing trend of global CO2 emissions over the past three decades is a wake-up call that we cannot ignore. Climate change, driven by these emissions, threatens not only our environment but also our way of life. This analysis dives into the data from 1990 to 2020, shining a light on the major culprits behind this rise.

Our analysis shows that the energy sector stands out as the largest contributor to CO2 emissions, followed closely by transportation, industry, and residential sources. Despite a growing global awareness and discussions about climate action, the data paints a sobering picture: we are not making the significant progress needed to curb these emissions.

The message is clear; we need immediate, targeted policy changes to tackle this pressing issue. The Global Carbon Project has reiterated that time is of the essence. If we are to meet our climate goals and secure a sustainable future for generations to come, we must act now. Together, we can shift the narrative and drive the change necessary to protect our planet.

### References

1.  [Global Carbon Project (2003) Science Framework and Implementation. Earth System Science Partnership (IGBP, IHDP, WCRP, DIVERSITAS) Report No. 1; Global Carbon Project Report No. 1, 69pp, Canberra.](https://www.globalcarbonproject.org/global/pdf/reports/GCP_Report%20No.1_English.pdf)
